---
title: "Reproducible Research: Peer Assessment 1"
author: "Trevor Morgan"
date: "January 18, 2015"
output: html_document
---

###Introduction:
I will be using a sample dataset from a forked github repository.  This file is in the working directory of that the following analysis runs from.  The analysis and write-up is done in an R markdown file that is knit into a webpage.

## Loading and preprocessing the data
```{r}
setwd("/Users/mm57647/Desktop/Coursera_test/5-Reproducible_Research/Week2/Peer_Assessment1/RepData_PeerAssessment1")
activity <- read.csv("activity.csv", colClasses = c("numeric", "character", "numeric"))
head(activity)
```

## What is mean total number of steps taken per day?
```{r}
#####Remove the "NA" values
step_data <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
#####Make a Histogram
hist(step_data$steps, main = "Total Steps Each Day", xlab = "Steps per Day", ylab = "Number of Days", col = "blue")
#####Compute the mean and median
day_mean <- mean(step_data$steps)
day_mean
day_median <- median(step_data$steps)
day_median
```
## What is the average daily activity pattern?
```{r}
average_interval <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot(average_interval$interval, average_interval$steps, type="l", xlab="Interval", ylab="Number of Steps")
max_interval <- average_interval[which.max(average_interval$steps),1]
max_interval
```

## Imputing missing values
```{r}
#############Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
sum(!complete.cases(activity))
#############Devise a strategy for filling in all of the missing values in the dataset - use the average steps for that interval
imputed_data <- transform(activity,
                          steps = ifelse(is.na(activity$steps),
                                         average_interval$steps[match(activity$interval, average_interval$interval)],
                                         activity$steps
                                        )
                          )

imputed_step_data <- aggregate(steps ~ date, imputed_data, sum)
hist(imputed_step_data$steps, main = "Total Steps Each Day", xlab = "Steps per Day", ylab = "Number of Days", col = "blue")

day_mean <- mean(imputed_step_data$steps)
day_mean
day_median <- median(imputed_step_data$steps)
day_median
```

The mean and median are now equal.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
imputed_data$dayType <- as.Date(imputed_data$date)
imputed_data$dayType <- ifelse(weekdays(imputed_data$dayType) %in%  c("Saturday", "Sunday"),'weekend','weekday')


library(ggplot2)
base_ggplot <- qplot(x=interval, y=steps, data=imputed_data, geom='smooth', stat='summary', fun.y=mean)

complete_plot <- base_ggplot +
                 facet_grid(dayType ~ .) +
                 facet_wrap(~dayType,nrow=2) +
                 theme(strip.background = element_rect(fill="#ffe5cc")) +
                 labs(title='Average steps per day (weekdays vs weekends')
print(complete_plot)
```

There is a higher peak early on the weekdays most likely for work/school starting, and more overall activity throughout the day on weekends.
